Four Models for Severe Weather Deaths

### 2017-05-11 paper.Rnw
under length! Yay!

\setlength{\marginparwidth}{2cm}
\begin{document}
<<imports, cache=FALSE, echo=FALSE>>= 
read_chunk('zinb_sampler.R')
options(scipen=4, digits=3)
library('truncnorm')
@
<<dataset, echo=FALSE, cache=FALSE>>= 
load('small_storms.RData')
storm_details <- small_storms
@
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
We chose to use a subset of that data covering the years 1996 through 2016. The data coms from the NOAA website \cite{dataset}. For this project we focused on the deaths directly attributable to a type of weather event. In the interest of time, we considered two types of severe weather, Tornados and Flash Floods, reponsible for the second and third most direct deaths among weather types. Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts. In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each. The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models. Section~\ref{s:results} describes the results of the analysis. And Section~\ref{s:discussion} contains the conclusions.

\section{Methods}
\label{s:methods}
The deaths attributed to a severe weather event is `count' data. 
The most common model used for count data is the Poisson distribution. 
However for some weather events, the negative binomial model is a better fit, because the Poisson distribution assumes that the events being counted occur independently.

Fortunately, the vast majority of severe weather events in the United States involve no deaths, therefore we wanted to also account for the possibility of structural zeros, therefore we also considered zero inflated variants. 
These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$. We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.

\subsection{Poisson}
\label{ss:mPoisson}
The Poisson model has one parameter, $\lambda$ represents the expected number of occurances of the event of interest. For a single random variable $x$, the probability density is:
$$p(x|\lambda)=\frac{\lambda^xe^{-\lambda}}{x!}.$$
We use Jeffrey's prior, and model 
$$\lambda|X\sim\mathcal{G}amma(n\bar{X}+1/2,n).$$ 
For a full derivation see~\ref{a:dP}.

@@

\subsection{Zero Inflated Negative Binomial}
\label{ss:mZiNBinom}
The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
For a single $X$ the probability density is: 
$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r.$$
We take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. For a full derivation, see \ref{a:dZINB}. My posterior is:
$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)$$
This distribution does not factor nicely, so I will use the Metropolis algorithm to sample from it. Because this posterior does not suggest any obvious proposal distributions I will sample each iteratively from a truncated normal distribution centered at $\theta^*$, and with a variance that is tuned to yield an appropriate acceptance rate.

@@

Because the posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, we were able to sample from this distribution directly. For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $\lambda$. The results are summarized in Table~\ref{t:rPOIS}, while density plots can be found in Figure~\ref{f:poisnbdensity}.

\begin{table}
    \centering
    \caption{Posterior distributions of $\lambda$, for both event types. The second column relays the median of the sample for $\lambda$, with the 95 percent credible interval in parentheses.}
    \label{t:rPOIS}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$  \\ \hline
    Tornado & 0.0592 (0.0564, 0.0620) \\ \hline
    Flash Flood & 0.0179 (0.0170, 0.0189) \\ \hline
    \end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics[width=.45\textwidth]{figure/POIS_Density.pdf}
\includegraphics[width=.45\textwidth]{figure/NB_Density.pdf}
\caption{Posterior density plots for Poisson (left) and Negative Binomial (right) for flood parameters (in blue), and tornado parameters (black). The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
\label{f:poisnbdensity}
\end{figure}

@@

As in the case of the Poisson distribution, the conditional posterior from our Negative Binomial likelihood and noninformative prior had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly. Meanwhile we took $r$ as known. For both Tornados, and Flash Floods, 10,000 samples were taken, to simulate the variable of interest, $p$. The results are summarized in Table~\ref{t:rNB}, while density plots can be found in Figure~\ref{f:poisnbdensity}.

\begin{table}
    \centering
    \caption{Posterior distributions of $p$, for both event types. The second column relays the median of the sample for $p$, with the 95 percent credible interval in parentheses.}
    \label{t:rNB}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $p$  \\ \hline
    Tornado & 0.0559 (0.0534, 0.0585) \\ \hline
    Flash Flood & 0.0176 (0.0167, 0.0186) \\ \hline
    \end{tabular}
\end{table}

\subsection{Zero Inflated Poisson}
\label{ss:rZiPoisson}

In tuning the parameters of the two models, slightly different values were chosen for the two different event types. The proposal distribution selected for $\lambda$ was a gamma(2, 2) for tornados, and a gamma(1, 2) for flash floods. The proposal for p within the tornado model was a beta(1940, 60), whereas it was a beta(2945, 55) in the flash flood model.

For each variable of interest, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are summarized in Table~\ref{t:rZIP}. Density plots can be found in Figure~\ref{f:zipfig}.

\begin{table}
    \centering
    \caption{The posterior distributions of $\lambda$ and p, for both event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p.}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}

@@

\subsection{Zero Inflated Negative Binomial}
\label{ss:rZiNBinom}
<<zinbsampler, echo=FALSE>>=
@
<<densityplot, echo=FALSE>>=
@
<<zinbparams, echo=FALSE, cache=TRUE>>=
samplesize <- 1030000
burnin <- 30000
thin <- 100
r.var=.00002
p.var=.00001
sigma.var=.001
@
The ZINB model was fit using the Metropolis-Hastings algorithm with a multivariate truncated normal distribution used as the proposal distribution. For the source code of the sampler, please see \ref{a:cZINB}.

This model suffered severely from the curse of dimensionality. It was very slow to converge, and the very small variance needed for a reasonable acceptance led to high autocorrelation, meaning agressive thinning was necessary. %Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}. 

@@

In the end $\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, resulting in a final sample size of $\Sexpr{length(flood.zinb$sigma)}$. %The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$. 
The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}, though due to questions of convergence the results should viewed with caution.

@@

The DIC for floods and tornadoes respectively were $\Sexpr{as.integer(flood.dic.zinb)}$ and $\Sexpr{as.integer(tornado.dic.zinb)}$.

@@

@@

\end{tabular}

@@

\begin{figure}
	\centering
<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=2.5, fig.width=6>>=
par(mfrow=c(1,3))
double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
                    expression(paste(sigma, ' Posterior Density ')), 
                    expression(sigma), 
                    expression(p*group("(",paste(sigma, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$p, tornado.zinb$p,
                    expression(paste(p, ' Posterior Density ')), 
                    expression(p), 
                    expression(p*group("(",paste(p, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$r, tornado.zinb$r,
                    expression(paste(r, ' Posterior Density ')), 
                    expression(r), 
                    expression(p*group("(",paste(r, "|", X),")")),
                    col=c('blue', 'black'))
@
	\caption{Posterior density curves for each of the three variables (ZINB model). The flood posterior is shown in blue and the tornado posterior is shown in black.}
	\label{f:zinbdensity}
\end{figure} 
\section{Discussion}
\label{s:discussion}
The goal of this analysis was to determine which of the four models was the best fit for the data. In order to make this determination the DIC was calculated for each model on both tornados and flash floods. The results are summarized in Table~\ref{t:evalresults}.

\begin{table}
\centering
\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
\label{t:evalresults}
\begin{tabular}{lrr}
\toprule
Model & DIC Flash Flood & DIC Tornado\\
\midrule
Poisson & 14237 & 17606\\
Negative Binomial & 101548 & 26372\\
ZIP & 10454 & 8727 \\
ZINB & \Sexpr{as.integer(flood.dic.zinb)}&\Sexpr{as.integer(tornado.dic.zinb)}\\
\bottomrule
\end{tabular}
\end{table}

In general, models with smaller DIC are preferred to models with larger DIC. From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados. Thus for estimating the rate of leathality of Flash Floods, Tornados, and possibly other severe weather, we should use a Zero Inflated Poisson model to fit the data.

\clearpage
\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}

>>>>

\setlength{\marginparwidth}{2cm}
\begin{document}
<<imports, cache=FALSE, echo=FALSE>>= 
read_chunk('zinb_sampler.R')
options(scipen=4, digits=3)
library('truncnorm')
@
<<dataset, echo=FALSE, cache=FALSE>>= 
load('small_storms.RData')
storm_details <- small_storms
@
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
Tornados and Flash Floods are reponsible for the second and third most direct deaths among weather types in the United States. Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts. NOAA's website has a database of weather events and the death toll available on their website \cite{dataset}. For this study we looked at events from 1996-2016.

The death toll from a severe weather event is a form of `count' data. 
The most common model used for count data is the Poisson distribution. 
However for some natural events, the negative binomial model is a better fit, because it is more flexible and allows for a difference between mean and variance.

Fortunately, the vast majority of severe weather events in the United States involve no deaths, so to account for the possibility of structural zeros, therefore we also considered zero inflated variants. 
These distributions are created by returning $0$ with probability $\sigma$ and sampling from the original distribution with probability $(1-\sigma)$. We will derive and fit a model for each of the four distributions and see if there is a difference in our results and evaluate to determine which model best fits the data.

The remainder of this analysis is organized as follows: Section~\ref{s:methods} discusses and derives the models. Section~\ref{s:results} describes the results of the analysis. And Section~\ref{s:discussion} contains the conclusions.

\section{Methods}
\label{s:methods}
\subsection{Poisson}
\label{ss:mPoisson}
The Poisson model has one parameter, $\lambda$ represents the expected number of occurances of the event of interest. For a single random variable $x$, the probability density is:
$$p(x|\lambda)=\frac{\lambda^xe^{-\lambda}}{x!}.$$
We use Jeffrey's prior, and model 
$$\lambda|X\sim\mathcal{G}amma(n\bar{X}+1/2,n).$$ 
For a full derivation see~\ref{a:dP}.

@@

\subsection{Zero Inflated Negative Binomial}
\label{ss:mZiNBinom}
The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
For a single $X$ the probability density is: 
$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r.$$
We take the uniform priors for $\sigma$ and $p$ as well as the non-informative gamma for $r$ which is $r^{-1/2}$. For a full derivation, see \ref{a:dZINB}. My posterior is:
$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}r^{-1/2}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{\Gamma(r)}\right)$$
This distribution does not factor, so I will use the Metropolis-Hastings algorithm to sample from it. I will use a truncated normal distribution centered at $\theta^*$ proposal density, and tune the variance for efficient convergence.

@@

The posterior from our Poisson likelihood and noninformative prior had a recognizable kernel, namely $\mathcal{G}amma(n\bar{X}+1/2,n)$, therefore we were able to sample from this distribution directly. For both event types, 10,000 samples were taken, to simulate the variable of interest, $\lambda$. The results are summarized in Table~\ref{t:rPOIS}, while density plots can be found in Figure~\ref{f:poisnbdensity}.

\begin{table}
    \centering
    \caption{Posterior distributions of $\lambda$, for both event types. The second column relays the median of the sample for $\lambda$, with the 95 percent credible interval in parentheses.}
    \label{t:rPOIS}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type &  Tornado &  Flash Flood \\ \hline
    $\lambda$ & 0.0592 (0.0564, 0.0620)  & 0.0179 (0.0170, 0.0189)\\
    \hline
    \end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics[width=.4\textwidth]{figure/POIS_Density.pdf}
\includegraphics[width=.4\textwidth]{figure/NB_Density.pdf}
\caption{Posterior density plots for Poisson (left) and Negative Binomial (right) for flood parameters (in blue), and tornado parameters (black). The solid line represents the median and the dashed lines indicate the 95\% confidence interval.}
\label{f:poisnbdensity}
\end{figure}

@@

The conditional posterior from our Negative Binomial likelihood and noninformative prior also had a recognizable kernel for $p$, namely $\mathcal{B}eta(n\bar{X},nr+1/2)$, thus we were able to sample from this distribution directly. We took $r$ as known. For both event types, 10,000 samples were taken. The results are summarized in Table~\ref{t:rNB}, and density plots can be found in Figure~\ref{f:poisnbdensity}.
\subsection{Zero Inflated Poisson}
\label{ss:rZiPoisson}
The proposal distribution selected for $\lambda$ was a gamma(2, 2) for tornados, and a gamma(1, 2) for flash floods. The proposal for p within the tornado model was a beta(1940, 60), whereas it was a beta(2945, 55) in the flash flood model. For each variable, 20,000 samples were taken. As convergence was not immediately achieved, the first 10,000 samples were discarded as a burn-in. The results are summarized in Table~\ref{t:rZIP}. Density plots can be found in Figure~\ref{f:zipfig}.

\begin{table}
    \centering
      \caption{The posterior distributions of $\lambda$ and p, for both event types. The second column relays the mean and 95 percent credible interval in parentheses. The third does the same for p.}
      \label{t:rZIP}
      \begin{tabular}{| l | l | l |}
      \hline
      Event Type & $\lambda$ & p  \\ \hline
      Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
      Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
      \end{tabular}
\end{table}

@@

\subsection{Zero Inflated Negative Binomial}
\label{ss:rZiNBinom}
<<zinbsampler, echo=FALSE>>=
<<densityplot, echo=FALSE>>=
<<zinbparams, echo=FALSE, cache=TRUE>>=
samplesize <- 2030000
burnin <- 30000
thin <- 250
r.var=.00002
p.var=.00001
sigma.var=.001

@@

The ZINB model was fit using the Metropolis-Hastings algorithm with a truncated multivariate normal proposal distribution. For the source code of the sampler, please see \ref{a:cZINB}. Because it was fitting 3 parameters simultaniously this model was slow to converge, and had high autocorrelation, requiring agressive thinning. %Running mean plots the Flash Flood models can be found in Figure~\ref{f:zinbconvergence}. 
$\Sexpr{samplesize}$ samples were drawn for each model, with a burn in of $\Sexpr{burnin}$ and one of every $\Sexpr{thin}$ samples was retained, for a final sample size of $\Sexpr{length(flood.zinb$sigma)}$. %The flood model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$ and the tornado model had an acceptance rate of $\Sexpr{mean(flood.zinb$ar)}$. 
The resulting densities are pictured in Figure~\ref{f:zinbdensity}, and Table~\ref{t:ZINBparams}. The densities for $p$ and $r$ are readily interpretable:  $p$ can be understood to measure dangerousness and $r$ can be thought of as duration. This implies floods are longer lasting than tornadoes, but less dangerous.

@@

@@

\scalebox{.8}{

@@

\end{tabular}}

@@

\begin{figure}
	\centering\scalebox{.8}{
<<zinbdensity, cache=TRUE, echo=FALSE, dependson=c('floodsample', 'tornadosample'), fig.height=2.5, fig.width=6>>=
par(mfrow=c(1,3))
double.density.plot(flood.zinb$sigma, tornado.zinb$sigma,
                    expression(paste(sigma, ' Posterior Density ')), 
                    expression(sigma), 
                    expression(p*group("(",paste(sigma, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$p, tornado.zinb$p,
                    expression(paste(p, ' Posterior Density ')), 
                    expression(p), 
                    expression(p*group("(",paste(p, "|", X),")")),
                    col=c('blue', 'black'))
double.density.plot(flood.zinb$r, tornado.zinb$r,
                    expression(paste(r, ' Posterior Density ')), 
                    expression(r), 
                    expression(p*group("(",paste(r, "|", X),")")),
                    col=c('blue', 'black'))
@
}
	\caption{Posterior density curves for each of the three variables (ZINB model). The flood posterior is shown in blue and the tornado posterior is shown in black.}
	\label{f:zinbdensity}
\end{figure} 
\section{Discussion}
\label{s:discussion}
\begin{table}
\centering
\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
\label{t:evalresults}
\scalebox{.8}{
\begin{tabular}{lrr}
\toprule
Model & DIC Flash Flood & DIC Tornado\\
\midrule
Poisson & 14237 & 17606\\
Negative Binomial & 101548 & 26372\\
ZIP & 10454 & 8727 \\
ZINB & \Sexpr{as.integer(flood.dic.zinb)}&\Sexpr{as.integer(tornado.dic.zinb)}\\
\bottomrule
\end{tabular}}
\end{table}
The goal of this analysis was to determine which of the four models best fit the data. In order to make this determination the DIC was calculated for each model on both tornados and flash floods. The results are summarized in Table~\ref{t:evalresults}.
The Zero Inflated Poisson model has the lowest DIC, for Flash Floods and the ZINB has the lowest for Tornados. For estimating or predicting lethality counts for weather it appears that zero inflated models out perform their plain variants however, which zero inflated model is not clear, as we only did two weather events and the ZINB and ZIP were similar. When attempting a regression model of death counts from weather events, researchers should consider a zero inflated model.

Negative binomial models tend to perform better than Poisson models when the mean is not equal to the varaince. It may be that this was true for the Tornadoes, but not true for the floods. The overdispersion of the Tornadoes can be clearly seen when comparing the posterior $\lambda$s for the ZIP model in Figure~\ref{f:zipfig}.

\begin{thebibliography}{2}
  \bibitem{dataset} NOAA's Severe Weather Data Inventory, 
    \url{https://www1.ncdc.noaa.gov/pub/data/swdi/stormevents/csvfiles/}. 
    Accessed April 2017.
  \bibitem{ZIP} Bayarri, M., Berger, J., Datta, G. (2008). Objective testing of Poisson versus inflated Poisson models. IMS 	Collections, 3, 105-121. 
\end{thebibliography}

### 2017-05-11 presentation.tex
under length! Yay!

These are unrecognizable. I ended up alternatively sampling from the two using Metropolis Hastings. For lambda I settled on a gamma proposal, and for p I settled on a Beta proposal. 
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		Convergence of parameters (flood model):
		\centering
		\includegraphics[width=.9\textwidth]{figure/ZIP_Flood_Conv.pdf} 
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		\begin{table}
    \centering
    \caption{After taking 20,000 samples from each variable and discarding 10,000 as a burn-in, we obtained the following (values are the mean of the posterior sample followed by 95 percent credible interval in parentheses)}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		Posterior Densities (Blue = Flood; Black = Tornado):
		\centering
		\includegraphics[width=.9\textwidth]{figure/ZIP_Density.pdf} 
	\end{frame}	
	\section{Zero Inflated Negative Binomial Model}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
		For a single $X$ the probability density is: 
		$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r$$
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		We took the uniform priors on $\sigma$ and $p$ and the non-informative prior $r^{-1/2}$ on $r$.
		This gave a posterior of:
		\scriptsize$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{r^{1/2}\Gamma(r)}\right)$$
		\normalsize This factors into nothing. Therefore we used the Metropolis Hastings Algorithm with proposal distributions of truncated normals.
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model:Results}
		\begin{itemize}
			\item Curse of Dimensionality; very slow convergence.
			\item High autocorrelation
			\item Low confidence in results
		\end{itemize}
		\begin{table}
			\centering
			\scalebox{0.6}{
			\begin{tabular}{lcccccc}
				\toprule
				&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
				\cmidrule(r){2-4}\cmidrule(l){5-7}
				Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
				\midrule
				$\sigma$ & 0.307 &
				0.282&
				(0.013, 
				0.716)&
				0.299 &
				0.276&
				(0.012, 
				0.702)\\
				$p$ &  0.554 &
				0.555&
				(0.518, 
				0.589)&
				0.862 &
				0.863&
				(0.841, 
				0.882)\\
				$r$ &  0.024 &
				0.02&
				(0.014, 
				0.052)&
				0.015 &
				0.013&
				(0.009, 
				0.033)\\
				\bottomrule
			\end{tabular}}
		\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		\begin{figure}[p]
			\centering
			\includegraphics[width=.9\textwidth]{figure/zinbdensity-1} 
			\caption{Posterior density curves for each of the three variables. The flood posterior is shown in blue and the tornado posterior is shown in black.}
		\end{figure} 
	\end{frame}

	\section{Conclusions}
	\begin{frame}
		\frametitle{Conclusions: Model Evaluation Table}
		% Waic dic table goes here
		\begin{table}
\centering
\caption{The following table shows the resulting DIC for each of the four models on both the tornado data and the flash flood data.}
\label{t:evalresults}
\begin{tabular}{lcc}
\toprule
Model & DIC Flash Flood & DIC Tornado\\
\midrule
Poisson & 14237 & 17606\\
Negative Binomial & 101548 & 26372\\
ZIP & 10454 & 8727 \\
ZINB & 10804.104&7166.309\\
\bottomrule
\end{tabular}
\end{table}

Models with smaller DIC are preferred. From the table we can see that the Zero Inflated Poisson model has the lowest DIC, both for Flash Floods and Tornados. Thus for estimating the rate of lethality of Flash Floods, Tornados, we should use a Zero Inflated Poisson model.
	\end{frame}
	\begin{frame}

		\frametitle{Conclusions: Future Work}
		% List of things we might have done if we'd had time
		\begin{enumerate}
			\item{Re-parameterize the ZINB model/look into other techniques to improve convergence.}
			\item{See if the results hold on other extreme weather events.}
			\item{Attempt to fit all data with a mixture model.}
		\end{enumerate}
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions}
		Questions?
	\end{frame}
\end{document}

>>>>

These are unrecognizable. I ended up alternatively sampling from the two using Metropolis Hastings. For lambda I settled on a gamma proposal, and for p I settled on a Beta proposal. 
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		Convergence of parameters (flood model):
		\centering
		\includegraphics[width=.9\textwidth]{figure/ZIP_Flood_Conv.pdf} 
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		\begin{table}
    \centering
    \caption{After taking 20,000 samples from each variable and discarding 10,000 as a burn-in, we obtained the following (values are the mean of the posterior sample followed by 95 percent credible interval in parentheses)}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Poisson Model}
		Posterior Densities (Blue = Flood; Black = Tornado):
		\centering
		\includegraphics[width=.9\textwidth]{figure/ZIP_Density.pdf} 
	\end{frame}	
	\section{Zero Inflated Negative Binomial Model}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		The Zero Inflated Negative Binomial (ZINB) model has three parameters $\sigma,$ the probability of a structural zero, and $p,r$ the usual negative binomial parameters. 
		For a single $X$ the probability density is: 
		$$p(X|\sigma, p, r) = \sigma I_{X=0}(X) + (1-\sigma)\frac{\Gamma(r+X)}{\Gamma(r)X!}p^X(1-p)^r$$
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		We took the uniform priors on $\sigma$ and $p$ and the non-informative prior $r^{-1/2}$ on $r$.
		This gave a posterior of:
		\scriptsize$$p(r,\sigma, p|X)\propto\left(\sigma + (1-\sigma)(1-p)^r\right)^Z(1-\sigma)^{N-Z}(1-p)^{(N-Z)r}p^{\sum_{i=1}^NX_i}\prod_{i=1}^N\left(\frac{\Gamma(r+X_i)}{r^{1/2}\Gamma(r)}\right)$$
		\normalsize This factors into nothing. Therefore we used the Metropolis Hastings Algorithm with proposal distributions of truncated normals.
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model:Results}
			$2030000$ samples were drawn for each model, with a burn in of $30000$ and one of every $250$ samples was retained, resulting in a final sample size of $8000$.
		\begin{itemize}
			\item Curse of Dimensionality; very slow convergence.
			\item High autocorrelation
			\item Low confidence in results
		\end{itemize}
		\begin{table}
			\centering
			\scalebox{.6}{
			\begin{tabular}{lcccccc}
				\toprule
				&\multicolumn{3}{c}{Flash Flood}&\multicolumn{3}{c}{Tornado}\\
				\cmidrule(r){2-4}\cmidrule(l){5-7}
				Parameter & Mean Value &Median Value & (95\% CI)& Mean Value &Median Value & (95\% CI)\\
				\midrule
				$\sigma$ & 0.308 &
				0.287&
				(0.012, 
				0.715)&
				0.29 &
				0.262&
				(0.012, 
				0.693)\\
				$p$ &  0.555 &
				0.555&
				(0.518, 
				0.591)&
				0.863 &
				0.863&
				(0.842, 
				0.882)\\
				$r$ &  0.024 &
				0.02&
				(0.014, 
				0.053)&
				0.015 &
				0.013&
				(0.009, 
				0.031)\\
				\bottomrule
			\end{tabular}}
		\end{table}
	\end{frame}
	\begin{frame}
		\frametitle{Zero Inflated Negative Binomial Model}
		\begin{figure}[p]
			\centering
			\includegraphics[width=.8\textwidth]{figure/zinbdensity-1} 
		\end{figure} 
		\begin{itemize}
			\item{$p$ can be understood to be ``dangerousness''}
			\item{$r$ can be thought of as ``duration''}
		\end{itemize}
		In this light results make sense: floods are longer lasting (on average) than tornadoes, but less dangerous (on average).
	\end{frame}

	\section{Conclusions}
	\begin{frame}
		\frametitle{Conclusions: Model Evaluation Table}
		% Waic dic table goes here
\begin{table}
	\centering
	\label{t:evalresults}
	\begin{tabular}{lrr}
		\toprule
		Model & DIC Flash Flood & DIC Tornado\\
		\midrule
		Poisson & 14237 & 17606\\
		Negative Binomial & 101548 & 26372\\
		ZIP & 10454 & 8727 \\
		ZINB & 10804&7166\\
		\bottomrule
	\end{tabular}
\end{table}

The Zero Inflated Poisson model has the lowest DIC, for Flash Floods and the ZINB has the lowest for Tornadoes. For estimating or predicting lethality counts for weather it appears that Zero inflated models out perform their plain variants however, which zero inflated model is not clear, as we only did two weather events and the ZINB and ZIP were similar.
	\end{frame}
	\begin{frame}

		\frametitle{Conclusions: Future Work}
		% List of things we might have done if we'd had time
		\begin{enumerate}
			\item{Re-parameterize the ZINB model/look into other techniques to improve convergence.}
			\item{See if the results hold on other extreme weather events.}
		\end{enumerate}
	\end{frame}
	\begin{frame}
		\frametitle{Conclusions}
		Questions?
	\end{frame}
\end{document}
